---
title: "test 9.2"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### Load packages
```{r}
library(leaflet)
library(dplyr)
library(tidyr)
library(ggplot2)
library(DT)
library(scales)
```

### Import data from KNB
```{r}
data_url <- "https://knb.ecoinformatics.org/knb/d1/mn/v2/object/urn%3Auuid%3Af119a05b-bbe7-4aea-93c6-85434dcb1c5e"

esc <- tryCatch(
  read.csv("data/escapement.csv", stringsAsFactors = FALSE),
  error = function(cond) {
    message(paste("Escapement file does not seem to exist, so get it from the KNB."))
    esc <-
      read.csv(url(data_url, method = "libcurl"), stringsAsFactors = FALSE)
    return(esc)
  }
)

head(esc)
```

### Data wrangling: calculate annual escapement by species and region
```{r}
annual_esc <- esc %>% 
  separate(sampleDate, c("Year", "Month", "Day"), sep = "-") %>% 
  mutate(Year = as.numeric(Year)) %>% 
  group_by(Species, SASAP.Region, Year) %>% 
  summarize(escapement = sum(DailyCount)) %>% 
  filter(Species %in% c("Chinook", "Sockeye", "Chum", "Coho", "Pink"))

head(annual_esc)
```

### Create a static figure using `ggplot2`
```{r}
ggplot(annual_esc, aes(x = Species, y = escapement, fill = SASAP.Region)) +
  geom_col()
```

### An aside: piping data into ggplot
```{r}
annual_esc %>%
  filter(SASAP.Region == "Kodiak") %>%
  ggplot(aes(x = Year, y = escapement, color = Species)) +
  geom_line() +
  geom_point()
```

### Back to creating static figures -- separate data wrangling & plotting sections of the above code and play with adjusting the theme
```{r}
kodiak_esc <- annual_esc %>%
  filter(SASAP.Region == "Kodiak") 
```

```{r}
ggplot(kodiak_esc, aes(x = Year, y = escapement, color = Species)) +
  geom_line() +
  geom_point() +
  ylab("Escapement") +
  ggtitle("Kodiak Salmon Escapement") +
  theme_bw() +
  theme(legend.position = "bottom", legend.title = element_blank())
```

### You can save your theme and add it to any future plots too
```{r}
my_theme <- theme_bw() +
  theme(legend.position = "bottom", legend.title = element_blank())

ggplot(kodiak_esc, aes(x = Year, y = escapement, color = Species)) +
  geom_line() +
  geom_point() +
  ylab("Escapement") +
  ggtitle("Kodiak Salmon Escapement") +
  my_theme
```

### Rotate x-axis tick marks to a 45 degree angle
```{r}
ggplot(kodiak_esc, aes(x = Year, y = escapement, color = Species)) +
  geom_line() +
  geom_point() +
  ylab("Escapement") +
  ggtitle("Kodiak Salmon Escapement") +
  theme_bw() +
  theme(
    legend.position = "bottom",
    legend.title = element_blank(),
    axis.text.x = element_text(angle = 45)
  )
```

### Smarter tick labels using `scales`
```{r}
ggplot(kodiak_esc, aes(x = Year, y = escapement, color = Species)) +
  geom_line() +
  geom_point() +
  scale_y_continuous(labels = comma) + # scientific to normal comma notation, yay!
  ylab("Escapement") +
  ggtitle("Kodiak Salmon Escapement") +
  my_theme
```

### Creating multiple plots
```{r}
ggplot(annual_esc, aes(x = Year, y = escapement, color = Species)) +
  geom_line() +
  geom_point() +
  scale_y_continuous(labels = comma) + # scientific to normal comma notation, yay!
  facet_wrap(~SASAP.Region, scales = "free_y", ncol = 2) +
  ylab("Escapement") +
  ggtitle("Kodiak Salmon Escapement") +
  my_theme
```

### Create an interactive table of unique sampling locations using `DT`
```{r}
locations <- esc %>% 
  distinct(Location, Latitude, Longitude) %>% 
  drop_na()

datatable(locations)
```

### Create an interactive map using `leaflet`
```{r}
leaflet(locations) %>% 
  addTiles() %>% 
  addMarkers(lng = ~Longitude, lat = ~Latitude, popup = ~Location)
```

### Use `leaflet` to import Web Map Service tiles (this is an example using GEBCO WMS tiles)
```{r}
leaflet(locations) %>% 
  addWMSTiles("https://www.gebco.net/data_and_products/gebco_web_services/web_map_service/mapserv?",
              layers = "GEBCO_LATEST",
              attribution = "Imager reproduced from the GEBCO_2014 Grid, version 20150318, www.gebco.net") %>% 
  addCircleMarkers(lng = ~Longitude,
                  lat = ~Latitude,
                  popup = ~Location,
                  radius = 5,
                  # set fill properties
                  fillColor = "salmon",
                  fillOpacity = 1,
                  # set stroke properties
                  stroke = T,
                  weight = 0.5, color = "white", 
                  opacity = 1)
```

